Hybrid MILP to efficiently and accuratly \\ solve hard DNN verification instances

### 2024-09-17 ICLR25/formula.tex

\begin{definition}
	Let $v(b)$ be the value of $b$ in the LP solution of upper bound of $a$, and $v(\hat{b})$ be the value of $\hat{b}$.	Then the formula to estimate improvement of upper bound of $b$ is: \begin{align*}
		I_b \approx W_{ba}(v(\hat{b})-\ReLU(v(b))).
	\end{align*}

@@

\begin{definition} \label{2layer}
	Suppose we have a fixed target node $a$ and a fixed source node $b$ in two layers before $a$. Let $\UB,\LB$ be the function of precomputed upper and lower bound before the layer of $a$, and let $v$ be the function of values of nodes in the LP solution of upper bound of $a$. 
	To compute the improvement of of upper bound of $a$ by $b$, we define the following function for nodes $c$ in one layer before $a$:
	\begin{align*}
		k(c) =
		\begin{cases}
			1, & \text{if } v(c) > 0 \text{ and } W_{ca} < 0\\
			0, & \text{if } v(c) \leq 0 \text{ and } W_{ca} < 0\\
			\frac{\UB(c)}{\UB(c)-\LB(c)}, & \text{if }W_{ca} > 0
		\end{cases}
	\end{align*} 
\end{definition}

We may use the following formula to estimate the improvement:  \begin{align*}
	I_b \approx (v(\hat{b})-\ReLU(v(b)))\sum_c W_{bc}W_{ca}k(c).
\end{align*}However, the error is still too large. This is because the improvement of a node $c$ can not exceed its precomputed upper and lower bound. Considering this, we may improve our formula.

\begin{definition}
	Continue the assumption in Definition \ref{2layer} To compute the improvement of of upper bound of $a$ by $b$, we define a distance function $D$:
	\begin{align*}
		&D(b) = \ReLU(v(b))-v(\hat{b})\\
			&D(c) =
		\begin{cases}
			\max(\LB(c), 0, v(c)+D(b)k(c)W_{bc})-v(\hat{c}), & \text{if }W_{ca} < 0\\
			\min(\UB(c), v(\hat{c})+D(b)k(c)W_{bc})-v(\hat{c}), & \text{if }  W_{ca} > 0
		\end{cases}\\
		&D(a) = \sum_c D(c)W_{ca}
	\end{align*}
\end{definition}

@@

This formula is based on previous subsection but more complex. In some network, run this formula may cost too much time but obtain little accuracy.

The key problem is how to compute the coefficient $k$ for nodes in two layers before the target layer. We use the following method to estimate the coefficient.

\begin{definition}\label{3layer}
	For a fixed target node $a$ and a fixed source node $b$ in two layers before $a$. Let $\UB$ and $\LB$ denote the precomputed upper bounds and lower bound used in building MILP models. To compute the improvement of of upper bound of $a$ by $b$, let $v$ be the function of solutions of LP model;then we define the following function $k$, so that for a node $c$ in 2 layers before the target node $a$, the value $k(c)$ is computed as:
	\begin{align*}
		&v_0 = v(\hat{c}), v_1 = \ReLU(v(c)), v_2 = \frac{\UB(c)v(c)-\UB(c)\LB(c)}{\UB(c)-\LB(c)}\\
		&k(c) =
		\begin{cases}
			\frac{v_0-v_1}{v_2-v_1}, & \text{if } v_2-v_1 > 0\\
			0.5, & \text{otherwise.}
		\end{cases}
	\end{align*} 
\end{definition}

@@

For $c$ in the next layer, we define \begin{align*}
	&u_0 = \max(\LB(c),\min(\UB(c),  v(c)+D(b)W_{bc}))\\
	&u_1 = \begin{cases}
		\ReLU(u_0)+k(c)(\frac{\UB(c)u_0-\UB(c)\LB(c)}{\UB(c)-\LB(c)}-u_0), & \text{if }\LB(c) < 0\\
	u_0, & \text{if }  \LB(c) \geq 0
	\end{cases}\\
	&D(c) = u_1-v(\hat{c})
\end{align*}

	Next is to compute $D(d)$ for nodes $d$ in one layer before $a$.

	\begin{align*}
		&w_0 = \sum_C D(c)W_{cd}\\
		&w_1 = \min(\UB(d),v(d)+w_0)\\		
		&D(d) =
		\begin{cases}
			w_1-v({d}), & \text{if }W_{da} > 0 \text{ and } \LB(d)\geq 0\\
		k(d)(w_1-v({d})), & \text{if }W_{da} > 0 \text{ and } \LB(d)< 0\\
		\ReLU(w_1)-v(\hat{d})	, & \text{if }  W_{da} < 0
		\end{cases}\\
		&D(a) = \sum_c D(d)W_{da}
	\end{align*}
\end{definition} $-D(a)$ is the estimation of improvement of upper bound of $a$ by opening node $b$.  For the lower bound, we can compute $D(a)$ by the same formula with the $v$ for lower bound LP solution and negative weights  instead.

>>>>

\begin{definition}
	Let $v(b)$ be the value of $b$ in the LP solution of upper bound of $a$, and $v(\hat{b})$ be the value of $\hat{b}$.	Then the formula to estimate improvement of upper bound of $b$ is: \begin{align}
		I_b \approx W_{ba}(v(\hat{b})-\ReLU(v(b))).
	\end{align}

@@

\begin{definition} \label{2layer}
	Suppose we have a fixed target node $a$ and a fixed source node $b$ in two layers before $a$. Let $\UB,\LB$ be the function of precomputed upper and lower bound before the layer of $a$, and let $v$ be the function of values of nodes in the LP solution of upper bound of $a$. 
	To compute the improvement of of upper bound of $a$ by $b$, we define the following function for nodes $c$ in one layer before $a$:
	\begin{align}
		k(c) =
		\begin{cases}
			1, & \text{if } v(c) > 0 \text{ and } W_{ca} < 0\\
			0, & \text{if } v(c) \leq 0 \text{ and } W_{ca} < 0\\
			\frac{\UB(c)}{\UB(c)-\LB(c)}, & \text{if }W_{ca} > 0
		\end{cases}
	\end{align} 
\end{definition}

We may use the following formula to estimate the improvement:  \begin{align*}
	I_b \approx (v(\hat{b})-\ReLU(v(b)))\sum_c W_{bc}W_{ca}k(c).
\end{align*}However, the error is still too large. This is because the improved value of a node $c$ can not exceed its precomputed upper and lower bound. Considering this, we can update our formula.

\begin{definition}
	Continue the assumption in Definition \ref{2layer} To compute the improvement of of upper bound of $a$ by $b$, we define a distance function $D$:
	\begin{align}
		&D(b) = \ReLU(v(b))-v(\hat{b})\\
			&D(c) =
		\begin{cases}
			\max(\LB(c), 0, v(c)+D(b)k(c)W_{bc})-v(\hat{c}), & \text{if }W_{ca} < 0\\
			\min(\UB(c), v(\hat{c})+D(b)k(c)W_{bc})-v(\hat{c}), & \text{if }  W_{ca} > 0
		\end{cases}\\
		&D(a) = \sum_c D(c)W_{ca}
	\end{align}
\end{definition}

@@

This formula is based on previous subsection but more complex. In some network, running this formula may cost too much time but obtain little accuracy.

The key problem is how to compute the coefficient $k$ for nodes in two layers before the target layer. To do this, we may use the values in the solution of LP model as follows:

\begin{definition}\label{3layer}
	For a fixed target node $a$ and a fixed source node $b$ in three layers before $a$. Let $\UB$ and $\LB$ denote the precomputed upper bounds and lower bound used in building MILP models. To compute the improvement of of upper bound of $a$ by $b$, let $v$ be the function of solutions of LP model; then we define the following function $h$ for all nodes $c$ in 2 layers before the target node $a$ as follows:
	\begin{align}
		&v_0 = v(\hat{c}), v_1 = \ReLU(v(c)), v_2 = \frac{\UB(c)v(c)-\UB(c)\LB(c)}{\UB(c)-\LB(c)}\\
		&h(c) =
		\begin{cases}
			\frac{v_0-v_1}{v_2-v_1}, & \text{if } v_2-v_1 > 0\\
			0.5, & \text{otherwise.}
		\end{cases}
	\end{align} 
\end{definition}

@@

To compute $D(c)$ for nodes $c$ in two layer before $a$, we define \begin{align}
	&u_0 = \max(\LB(c),\min(\UB(c),  v(c)+D(b)W_{bc}))\\
	&u_1 = \begin{cases}
		\ReLU(u_0)+h(c)(\frac{\UB(c)u_0-\UB(c)\LB(c)}{\UB(c)-\LB(c)}-\ReLU(u_0)), & \text{if }\LB(c) < 0\\
	u_0, & \text{if }  \LB(c) \geq 0
	\end{cases}\\
	&D(c) = u_1-v(\hat{c})
\end{align}

	To compute $D(d)$ for nodes $d$ in one layer before $a$, we define 
	\begin{align}
		&w_0 = \sum_c D(c)W_{cd}\\
		&w_1 = \min(\UB(d),v(d)+w_0)\\		
		&D(d) =
		\begin{cases}
			w_1-v({d}), & \text{if }W_{da} > 0 \text{ and } \LB(d)\geq 0\\
		k(d)(w_1-v({d})), & \text{if }W_{da} > 0 \text{ and } \LB(d)< 0\\
		\ReLU(w_1)-v(\hat{d})	, & \text{if }  W_{da} < 0
		\end{cases}\\
		&D(a) = \sum_c D(d)W_{da}
	\end{align}
\end{definition} $-D(a)$ is the estimation of improvement of upper bound of $a$ by opening node $b$.  For the lower bound, we can compute $D(a)$ by the same formula with the $v$ for lower bound LP solution and negative weights  instead.
